High-Fidelity Counterfactual Explanations through Conformal Prediction

### 2023-04-19 notebooks/mnist.qmd
:cry:

```{julia}
# Data:
n_obs = 1000
counterfactual_data = load_mnist(n_obs)
X, y = CounterfactualExplanations.DataPreprocessing.unpack_data(counterfactual_data)
X = table(permutedims(X))
labels = counterfactual_data.output_encoder.labels
input_dim, n_obs = size(counterfactual_data.X)
n_digits = Int(sqrt(input_dim))
output_dim = length(unique(labels))
```

@@

```{julia}
# Model parameters:
epochs = 100
batch_size = Int(round(n_obs/10))
n_hidden = 32
activation = Flux.swish
builder = MLJFlux.@builder Flux.Chain(
    Dense(n_in, n_hidden),
    BatchNorm(n_hidden, activation),
    Dense(n_hidden, n_out),
    BatchNorm(n_out)
)
# builder = MLJFlux.MLP(
#     hidden=(
#         n_hidden,
#         n_hidden,
#         n_hidden,
#     ), 
#     σ=activation
# )
α = [1.0,1.0,1e-1]

@@

# Joint Energy Model:
𝒟x = Uniform(0,1)
𝒟y = Categorical(ones(output_dim) ./ output_dim)
sampler = ConditionalSampler(𝒟x, 𝒟y, input_size=(input_dim,), batch_size=batch_size)
jem = JointEnergyClassifier(
    sampler;
    builder=builder,
    batch_size=batch_size,
    finaliser=x -> x,
    loss=Flux.Losses.logitcrossentropy,
    jem_training_params=(α=α,verbosity=10,),
    sampling_steps=20,
    epochs=epochs,
)

@@

# Set up search:
factual_label = 2
x = reshape(counterfactual_data.X[:,rand(findall(predict_label(M, counterfactual_data).==factual_label))],input_dim,1)
target = 0
factual = predict_label(M, counterfactual_data, x)[1]
γ = 0.5
T = 100

>>>>

```{julia}
# Data:
n_obs = 10000
counterfactual_data = load_mnist(n_obs)
X, y = CounterfactualExplanations.DataPreprocessing.unpack_data(counterfactual_data)
X = table(permutedims(X))
labels = counterfactual_data.output_encoder.labels
input_dim, n_obs = size(counterfactual_data.X)
n_digits = Int(sqrt(input_dim))
output_dim = length(unique(labels))
```

@@

```{julia}
# Model parameters:
epochs = 100
batch_size = minimum([Int(round(n_obs/10)), 100])
n_hidden = 32
activation = Flux.relu
builder = MLJFlux.@builder Flux.Chain(
    Dense(n_in, n_hidden),
    BatchNorm(n_hidden, activation),
    Dense(n_hidden, n_out),
    BatchNorm(n_out)
)
# builder = MLJFlux.Short(n_hidden=n_hidden, dropout=0.2, σ=activation)
# builder = MLJFlux.MLP(
#     hidden=(
#         n_hidden,
#         n_hidden,
#         n_hidden,
#     ), 
#     σ=activation
# )
α = [1.0,1.0,5e-1]

@@

# Joint Energy Model:
𝒟x = Uniform(0,1)
𝒟y = Categorical(ones(output_dim) ./ output_dim)
sampler = ConditionalSampler(𝒟x, 𝒟y, input_size=(input_dim,), batch_size=batch_size)
jem = JointEnergyClassifier(
    sampler;
    builder=builder,
    batch_size=batch_size,
    finaliser=x -> x,
    loss=Flux.Losses.logitcrossentropy,
    jem_training_params=(
        α=α,verbosity=10,
        use_gen_loss=false,
        use_reg_loss=false,
    ),
    sampling_steps=20,
    epochs=epochs,
)

@@

# Set up search:
factual_label = 9
x = reshape(counterfactual_data.X[:,rand(findall(predict_label(M, counterfactual_data).==factual_label))],input_dim,1)
target = 4
factual = predict_label(M, counterfactual_data, x)[1]
γ = 0.5
T = 1
